Forward and adjoint simulations of seismic wave propagation on fully unstructured hexahedral meshes

### 2013-09-20 doc/GJI_2011_paper_about_the_code/paper_GJI_2011_about_the_code.tex
converted all the export2SESAME() names in CUBIT scripts to the new GEOCUBIT name, which is export2SPECFEM3D()

%% ------------------------------------------------------------------------ %%
%  TITLE
%% ------------------------------------------------------------------------ %%
\title[SPECFEM3D Version~2.0 `Sesame']{Forward and adjoint simulations of seismic wave propagation on fully unstructured hexahedral meshes}

@@

\begin{summary}
We present forward and adjoint
spectral-element simulations of coupled acoustic and (an)elastic seismic
wave propagation on fully unstructured hexahedral meshes.
Simulations benefit from recent advances in hexahedral meshing, load balancing
and software optimization.
Meshing may be accomplished using a mesh generation tool kit such as \href{http://cubit.sandia.gov}{CUBIT},
and load balancing is facilitated by graph partitioning based on the \href{http://www.labri.fr/perso/pelegrin/scotch/}{SCOTCH} library.
Coupling between fluid and solid regions is incorporated in straightforward fashion using domain decomposition.
Topography, bathymetry and Moho undulations may be readily included in the mesh,
and physical dispersion and attenuation associated with anelasticity are accounted for using a series
of standard linear solids.
Finite-frequency Fr\'echet derivatives are calculated using adjoint methods in both fluid and solid domains.
The software is benchmarked for a layercake model.
We present various examples of fully unstructured meshes, snapshots of wavefields and finite-frequency kernels
generated by Version~2.0 `Sesame' of our widely used open source spectral-element package \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D}.
 \end{summary}

%% ------------------------------------------------------------------------ %%
\section{Introduction}\label{sec:Introduction}
%% ------------------------------------------------------------------------ %%
We present a new software package, \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame',
capable of simulating forward and adjoint seismic wave propagation on fully unstructured hexahedral meshes of
arbitrary shaped model domains.
In view of unrelenting  growth in computational power,
it has become more-and-more important to develop software capable of
harnessing powerful computers to address a broad range of seismological forward and inverse problems.
A well-established numerical technique for solving such problems in a fast and highly accurate manner
is the spectral-element method (SEM).
The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
and has been successfully adapted to address problems in seismic wave propagation.
Early seismic wave propagation applications of the SEM, utilizing Legendre basis functions and a
perfectly diagonal mass matrix, include \cite{CoJoTo93}, \cite{Kom97},
\cite{FaMaPaQu97}, \cite{CaGa97}, \cite{KoVi98} and \cite{KoTr99},
whereas applications involving Chebyshev basis functions and a nondiagonal mass matrix
include \cite{SePr94}, \cite{PrCaSe94} and \cite{SePrPr95}.

@@

The purpose of this article is to present forward and adjoint simulations in various 3D models using the new software package, \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame', thereby illustrating its current capabilities.
The original \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} package for local simulations was extended, improved and optimized in various ways.
The Version~2.0 `Sesame' release includes a more flexible internal mesher and accommodates more powerful external meshers,
such as \href{http://cubit.sandia.gov}{CUBIT}
\citep{Bl94,WhMiBe95,Mi96,CaStLeKoPiTr08}.
Adding such external meshers into the workflow greatly increases flexibility for high-performance applications,
as illustrated by the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software package \citep{CaGa97,StPaIg09,ChMoTsBaKrKaStKr10}.
Advantages of \href{http://geoelse.stru.polimi.it/}{GeoELSE} include the accommodation of visco-plastic and non-linear rheologies,
whereas benefits of \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} include coupled fluid-solid domains and
adjoint capabilities; the latter enable one to address seismological inverse problems.
Load balancing parallel simulations in \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} is accomplished based on the graph partitioning
software package \href{http://www.labri.fr/perso/pelegrin/scotch/}{SCOTCH}
\citep{PeRo96,ChPe08}.
The new package facilitates coupled forward and adjoint acoustic/(an)elastic simulations,
which are especially interesting for problems in exploration seismology, ocean acoustics and medical tomography.
The new software is freely available under the GNU GPL Version~2 license via \href{http://www.geodynamics.org}{CIG}.

@@

Let us briefly summarize the equations governing seismic wave propagation implemented in \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D}.
For more technical details, the reader is referred to \citet{KoTr99}.
\href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame' implements wave propagation in coupled (an)elastic and acoustic materials
on local scales.
We may thus safely neglect additional effects that would arise from self-gravitation and rotation
\citep{KoTr02b,KoTsTr05,ChKoViCaVaFe07}, which are important at longer periods.
In the following, we first discuss (an)elastic wave propagation and subsequently consider acoustic waves.

@@

We have taken advantage of recent advances in high-performance computing, fully unstructured hexahedral meshing, load balancing and
mesh partitioning to facilitate forward and adjoint simulations of seismic wave propagation in coupled fluid and solid domains.
Our new open source software package,
\href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame',
performs acoustic and (an)elastic simulations of seismic wave propagation in complex geological models.
Hexahedral meshes may be generated based on packages such as
\href{http://cubit.sandia.gov}{CUBIT},
\href{http://www.simulia.com/products/abaqus_fea.html}{Abaqus},
\href{http://www.ansys.com/}{ANSYS},
\href{http://www.gocad.org/www/}{GOCAD},
\href{http://gid.cimne.upc.es}{GiD},
\href{http://geuz.org/gmsh}{Gmsh},
\href{http://www.truegrid.com/}{TrueGrid}
or  \href{http://www.salome-platform.org/}{Salome},
but the simple in-house mesher used in previous versions of \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D}
remains available for back-compatibility.

@@

\begin{acknowledgments}
We thank two anonymous reviewers for comments which improved the manuscript.
All simulations were performed on a Dell cluster built and maintained by the
Princeton Institute for Computational Science \& Engineering (USA).
The open source spectral-element software package \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame' used
for this article is available for download via the Computational Infrastructure for Geodynamics
(\href{http://geodynamics.org/}{CIG}).
3D graphics were produced with the open source visualization application \href{http://www.paraview.org/}{Paraview}.
We acknowledge support by the US National Science Foundation under grant EAR-0711177.
\end{acknowledgments}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{./images/processing.jpg}
\end{center}
\caption{Workflow for running spectral-element simulations with \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 `Sesame'.
}
\label{figure:processing}
\end{figure}

>>>>

%% ------------------------------------------------------------------------ %%
%  TITLE
%% ------------------------------------------------------------------------ %%
\title[SPECFEM3D Version~2.0]{Forward and adjoint simulations of seismic wave propagation on fully unstructured hexahedral meshes}

@@

\begin{summary}
We present forward and adjoint
spectral-element simulations of coupled acoustic and (an)elastic seismic
wave propagation on fully unstructured hexahedral meshes.
Simulations benefit from recent advances in hexahedral meshing, load balancing
and software optimization.
Meshing may be accomplished using a mesh generation tool kit such as \href{http://cubit.sandia.gov}{CUBIT},
and load balancing is facilitated by graph partitioning based on the \href{http://www.labri.fr/perso/pelegrin/scotch/}{SCOTCH} library.
Coupling between fluid and solid regions is incorporated in straightforward fashion using domain decomposition.
Topography, bathymetry and Moho undulations may be readily included in the mesh,
and physical dispersion and attenuation associated with anelasticity are accounted for using a series
of standard linear solids.
Finite-frequency Fr\'echet derivatives are calculated using adjoint methods in both fluid and solid domains.
The software is benchmarked for a layercake model.
We present various examples of fully unstructured meshes, snapshots of wavefields and finite-frequency kernels
generated by Version~2.0 of our widely used open source spectral-element package \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D}.
 \end{summary}

%% ------------------------------------------------------------------------ %%
\section{Introduction}\label{sec:Introduction}
%% ------------------------------------------------------------------------ %%
We present a new software package, \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0,
capable of simulating forward and adjoint seismic wave propagation on fully unstructured hexahedral meshes of
arbitrary shaped model domains.
In view of unrelenting  growth in computational power,
it has become more-and-more important to develop software capable of
harnessing powerful computers to address a broad range of seismological forward and inverse problems.
A well-established numerical technique for solving such problems in a fast and highly accurate manner
is the spectral-element method (SEM).
The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
and has been successfully adapted to address problems in seismic wave propagation.
Early seismic wave propagation applications of the SEM, utilizing Legendre basis functions and a
perfectly diagonal mass matrix, include \cite{CoJoTo93}, \cite{Kom97},
\cite{FaMaPaQu97}, \cite{CaGa97}, \cite{KoVi98} and \cite{KoTr99},
whereas applications involving Chebyshev basis functions and a nondiagonal mass matrix
include \cite{SePr94}, \cite{PrCaSe94} and \cite{SePrPr95}.

@@

The purpose of this article is to present forward and adjoint simulations in various 3D models using the new software package, \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0, thereby illustrating its current capabilities.
The original \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} package for local simulations was extended, improved and optimized in various ways.
The Version~2.0 release includes a more flexible internal mesher and accommodates more powerful external meshers,
such as \href{http://cubit.sandia.gov}{CUBIT}
\citep{Bl94,WhMiBe95,Mi96,CaStLeKoPiTr08}.
Adding such external meshers into the workflow greatly increases flexibility for high-performance applications,
as illustrated by the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software package \citep{CaGa97,StPaIg09,ChMoTsBaKrKaStKr10}.
Advantages of \href{http://geoelse.stru.polimi.it/}{GeoELSE} include the accommodation of visco-plastic and non-linear rheologies,
whereas benefits of \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} include coupled fluid-solid domains and
adjoint capabilities; the latter enable one to address seismological inverse problems.
Load balancing parallel simulations in \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} is accomplished based on the graph partitioning
software package \href{http://www.labri.fr/perso/pelegrin/scotch/}{SCOTCH}
\citep{PeRo96,ChPe08}.
The new package facilitates coupled forward and adjoint acoustic/(an)elastic simulations,
which are especially interesting for problems in exploration seismology, ocean acoustics and medical tomography.
The new software is freely available under the GNU GPL Version~2 license via \href{http://www.geodynamics.org}{CIG}.

@@

Let us briefly summarize the equations governing seismic wave propagation implemented in \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D}.
For more technical details, the reader is referred to \citet{KoTr99}.
\href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 implements wave propagation in coupled (an)elastic and acoustic materials
on local scales.
We may thus safely neglect additional effects that would arise from self-gravitation and rotation
\citep{KoTr02b,KoTsTr05,ChKoViCaVaFe07}, which are important at longer periods.
In the following, we first discuss (an)elastic wave propagation and subsequently consider acoustic waves.

@@

We have taken advantage of recent advances in high-performance computing, fully unstructured hexahedral meshing, load balancing and
mesh partitioning to facilitate forward and adjoint simulations of seismic wave propagation in coupled fluid and solid domains.
Our new open source software package,
\href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0,
performs acoustic and (an)elastic simulations of seismic wave propagation in complex geological models.
Hexahedral meshes may be generated based on packages such as
\href{http://cubit.sandia.gov}{CUBIT},
\href{http://www.simulia.com/products/abaqus_fea.html}{Abaqus},
\href{http://www.ansys.com/}{ANSYS},
\href{http://www.gocad.org/www/}{GOCAD},
\href{http://gid.cimne.upc.es}{GiD},
\href{http://geuz.org/gmsh}{Gmsh},
\href{http://www.truegrid.com/}{TrueGrid}
or  \href{http://www.salome-platform.org/}{Salome},
but the simple in-house mesher used in previous versions of \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D}
remains available for back-compatibility.

@@

\begin{acknowledgments}
We thank two anonymous reviewers for comments which improved the manuscript.
All simulations were performed on a Dell cluster built and maintained by the
Princeton Institute for Computational Science \& Engineering (USA).
The open source spectral-element software package \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0 used
for this article is available for download via the Computational Infrastructure for Geodynamics
(\href{http://geodynamics.org/}{CIG}).
3D graphics were produced with the open source visualization application \href{http://www.paraview.org/}{Paraview}.
We acknowledge support by the US National Science Foundation under grant EAR-0711177.
\end{acknowledgments}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{./images/processing.jpg}
\end{center}
\caption{Workflow for running spectral-element simulations with \href{http://www.geodynamics.org/cig/software/specfem3d}{SPECFEM3D} Version~2.0.
}
\label{figure:processing}
\end{figure}
